---
title: "Plots"
author: "CGA"
output: html_document
---

## Libraries

```{r}
library(tidyverse)
library(viridis)
library(treemap)
library(vegan)
library(UpSetR)
library(hrbrthemes)
```

## Functions

### Taxonomy function (by Dr. Glasner)

Using this may result a bit hectic, but it will help you to understand how to navigate your data instead when others tools are too restrictive

```{r}
Tax.sum <- function(OTU.Table, Tax.Table, Tax.lvl ){
  z <- NULL
  y <- NULL
  for (i in 1:length(unique(Tax.Table[colnames(OTU.Table),Tax.lvl]))) {
    if (length(OTU.Table[,which(Tax.Table[colnames(OTU.Table),Tax.lvl]==unique(Tax.Table[colnames(OTU.Table),Tax.lvl])[i])])!=length(rownames(OTU.Table))) {
      z <- which(Tax.Table[colnames(OTU.Table),Tax.lvl]==unique(Tax.Table[colnames(OTU.Table),Tax.lvl])[i])
      y <- cbind(y, apply(OTU.Table[,which(Tax.Table[colnames(OTU.Table),Tax.lvl]==unique(Tax.Table[colnames(OTU.Table),Tax.lvl])[i])], 1, function(x) sum(x)))
    } else {
      y <- cbind(y, OTU.Table[,which(Tax.Table[colnames(OTU.Table),Tax.lvl]==unique(Tax.Table[colnames(OTU.Table),Tax.lvl])[i])])
    }
  }
  colnames(y) <- unique(Tax.Table[colnames(OTU.Table),Tax.lvl])
  invisible((y))
}
```

## Data import

```{r data}
autotrophics <- read.csv("https://raw.githubusercontent.com/cmlglvz/sentrynels/main/Data/PPE/Autotrophic_eukaryotes_ASV.csv", 
                         header = TRUE, 
                         sep = ";", 
                         skip = 0, 
                         row.names = 1)

ppes <- read.csv2("https://raw.githubusercontent.com/cmlglvz/sentrynels/main/Data/PPE/PPEs_ASV_table.csv", 
                  header = TRUE, 
                  sep = ";", 
                  skip = 0,
                  row.names = 1)
```

## Taxa level abundance

### Data preparation

Colors are based on Tragin & Vaulot (2018) publication. It will more relevant later (trust me)

```{r}
auto_taxa <- autotrophics[,-c(3:72)]
auto_abundance <- as.data.frame(t(autotrophics[,-c(1,2,66:81)]))
auto_phylum <- Tax.sum(auto_abundance, auto_taxa, 5) %>% 
  t() %>% 
  as.data.frame() %>% 
  tibble::add_column(Color = c("#90ED90", "#CDAE94", "#6CA6CD", "#7EFFD3", "#C577F9", "#F9A555", "#2E4756"))
auto_phylum <- auto_phylum %>%
  dplyr::mutate(Chanaral = rowSums(auto_phylum[c(1:12)]),
                Flamenco = rowSums(auto_phylum[c(13:24)]), 
                Huasco = rowSums(auto_phylum[c(25:36)]), 
                Pta_Choros = rowSums(auto_phylum[c(37:41)]),
                Quintero = rowSums(auto_phylum[c(42:51)]),
                Las_Cruces = rowSums(auto_phylum[c(52:63)]),
                Total = rowSums(auto_phylum[c(1:63)]), 
                .after = "L4A18")

ppe_taxa <- ppes[,-c(3:72)]
ppe_abundance <- as.data.frame(t(ppes[,-c(1,2,66:81)]))
ppe_phylum <- Tax.sum(ppe_abundance, ppe_taxa, 5) %>% 
  t() %>% 
  as.data.frame() %>% 
  tibble::add_column(Color = c("#90ED90", "#CDAE94", "#6CA6CD", "#7EFFD3", "#C577F9", "#2E4756"))
ppe_phylum <- ppe_phylum %>%
  dplyr::mutate(Chanaral = rowSums(ppe_phylum[c(1:12)]),
                Flamenco = rowSums(ppe_phylum[c(13:24)]), 
                Huasco = rowSums(ppe_phylum[c(25:36)]), 
                Pta_Choros = rowSums(ppe_phylum[c(37:41)]),
                Quintero = rowSums(ppe_phylum[c(42:51)]),
                Las_Cruces = rowSums(ppe_phylum[c(52:63)]),
                Total = rowSums(ppe_phylum[c(1:63)]), 
                .after = "L4A18")
```

### Treemap plots

```{r autotrophs}
png("/Users/Administrador/Desktop/Camilo/sentrynels/Analysis/PPEs/Fig_8_treemap_autotrophic_ASV_all_sites.png", width = 23, height = 18, units = "in", res = 600)
treemap(as_tibble(auto_phylum, rownames = "Phylum", .name_repair = "minimal"), # by doing this we avoid problems related to using row names as "Index"
        index = "Phylum", 
        vSize = "Total", # We can change this to get the distribution at each site
        type = "color", # Here we indicate how are we "painting" each section
        vColor = "Color", # And here we point at the colors we want. In this case we tell the function to look at the column we previously added with the colors  from the article
        position.legend = "none", 
        fontsize.labels = 55, 
        title = "Autotrophic eukaryotes ASV distribution (counts) for complete dataset", 
        fontsize.title = 42, 
        title.legend = NA, 
        border.col = NA)
dev.off()
```

```{r ppes}
png("/Users/Administrador/Desktop/Camilo/sentrynels/Analysis/PPEs/Fig_9_treemap_PPE_ASV_all_sites.png", width = 23, height = 18, units = "in", res = 600)
treemap(as_tibble(ppe_phylum, rownames = "Phylum", .name_repair = "minimal"), 
        index = "Phylum", 
        vSize = "Total", 
        type = "color", 
        vColor = "Color", 
        position.legend = "none", 
        fontsize.labels = 55, 
        title = "PPE ASVs distribution (counts) for complete dataset", 
        fontsize.title = 42, 
        title.legend = NA, 
        border.col = NA)
dev.off()
```

## Relative abundance

### Phylum - Autotrophs

```{r}
rel_abund_auto <- auto_phylum[,-c(64:71)] %>% 
  t() %>% 
  as.data.frame() %>% 
  add_column(Site = c(rep("Chanaral",12), 
                      rep("Flamenco",12), 
                      rep("Huasco",12),
                      rep("Pta_Choros",5),
                      rep("Quintero",10),
                      rep("Las_Cruces",12)), 
             .before = "Chlorophyta", 
             .name_repair = "minimal")
rel_abund_auto$Site <- factor(rel_abund_auto$Site, levels = unique(rel_abund_auto$Site))
long_raa <- reshape2::melt(rel_abund_auto, id.vars = "Site", variable.name = "Phylum") %>% dplyr::rename(Abundance = value)
inverted <- factor(long_raa$Phylum, levels = rev(as.character(unique(long_raa$Phylum))))
```

### Plot

```{r}
raaplot <- ggplot(long_raa,
                  aes(x = Site,
                      y = Abundance,
                      fill = inverted)) + 
  geom_bar(position = "fill", stat = "identity") + 
  labs(x = "Sites", 
       y = "Relative abundance", 
       title = "Relative abundance of autotrophic reads per site") + 
  scale_fill_manual(values = rev(as.character(auto_phylum$Color))) + 
  theme_ipsum(base_size = 22, 
              axis_title_size = 25, 
              plot_title_size = 25) + 
  theme(legend.position = "bottom", 
        legend.direction = "horizontal")
ggsave(plot = raaplot, 
       filename = "Fig_10_relative_reads_abundance_autotrophs_per_site.png", 
       path = "/Users/Administrador/Desktop/Camilo/sentrynels/Analysis/PPEs/", 
       width = 23, 
       height = 15, 
       dpi = 600)
```

### Class - PPEs

```{r}
ppe_class <- Tax.sum(ppe_abundance, ppe_taxa, 7) %>% as_tibble(rownames = "Samples", .name_repair = "minimal")
ppe_class$Samples <- factor(ppe_class$Samples, levels = unique(ppe_class$Samples))
long_ppeclass <- reshape2::melt(ppe_class[,-25], id.vars =  "Samples", variable.name = "Class") %>% rename(Abundance = value)
reversed <- rev(as.character(c("Mamiellophyceae", "Trebouxiophyceae", "Chlorodendrophyceae", "Chloropicophyceae", "Pyramimonadophyceae", "Prasino-Clade-9", "Chlorophyceae", "Nephroselmidophyceae", "Prasino-Clade-VIII", "Coscinodiscophyceae", "Pelagophyceae", "Mediophyceae", "Chrysophyceae", "Bolidophyceae", "Bacillariophyceae", "Dictyochophyceae", "MOCH-3", "MOCH-5", "MOCH-2", "MOCH-1", "Xanthophyceae", "MOCH-4", "Prymnesiophyceae", "Haptophyta_Clade_HAP3", "Haptophyta_Clade_HAP5", "Haptophyta_Clade_HAP4", "Cryptophyceae", "Cryptophyceae:nucl", "Kathablepharidea")))
long_ppeclass$Class <- factor(long_ppeclass$Class, levels = unique(reversed))
```

### Plot

```{r}
claplot <- ggplot(long_ppeclass,
                  aes(x = Samples,
                      y = Abundance,
                      fill = Class)) + 
  geom_bar(position = "fill", stat = "identity", width = 1, color = "darkgrey") + 
  labs(x = "Samples", 
       y = "Relative abundance", 
       title = "Relative abundance of PPE reads grouped by class at each sample") + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.05, end = 0.95, direction = -1, option = "turbo", alpha = 0.7) + 
  scale_fill_manual(values = c("#C577F9", "#5888A8", "#6CA6CD", "#570D16", "#6D101B", "#981626", "#C41D31", "#4E433A", "#564A40", "#5F5146", "#69594D", "#736255", "#8C7766", "#9A8370", "#A9907B", "#BA9E87", "#CDAE94", "#D2B59E", "#D6BDA7", "#DBC4B1", "#274127", "#345634", "#416C41", "#4F814F", "#5C975C", "#69AC69", "#76C276", "#83D783", "#90ED90")) + 
  #scale_fill_manual(values = c("#C577F9", "#5888A8", "#6CA6CD", "#570D16", "#6D101B", "#981626", "#C41D31", "#4E433A", "#564A40", "#5F5146", "#69594D", "#736255", "#8C7766", "#9A8370", "#A9907B", "#BA9E87", "#CDAE94", "#D2B59E", "#D6BDA7", "#DBC4B1", "#DFCBBB", "#E4D3C5", "#E8DACE", "#EDE2D8", "#F1E9E2", "#F6F1ED", "#FBF9F7", "#274127", "#345634", "#416C41", "#4F814F", "#5C975C", "#69AC69", "#76C276", "#83D783", "#90ED90", "#A3F0A3", "#BFF5BF")) + 
  theme_ipsum(base_size = 22, 
              axis_title_size = 25, 
              plot_title_size = 25) + 
  theme(legend.position = "bottom", 
        legend.direction = "horizontal")
ggsave(plot = claplot, 
       filename = "Fig_11_ver_2_relative_reads_abundance_ppes_by_class.png", 
       path = "/Users/Administrador/Desktop/Camilo/sentrynels/Analysis/PPEs/", 
       width = 35, 
       height = 20, 
       dpi = 600)
```

