---
title: "Phylo-like analysis"
author: "CGA"
output: html_document
---

## Libraries

```{r, echo=FALSE,message=FALSE}
library(tidyverse)
library(viridis)
library(vegan)
library(UpSetR)
library(hrbrthemes)
```

## Data import

```{r data}
previous <- read.csv("/Documents/Proyectos_R/sentrynels/Data/Phylo/Previous_core_ppe_abundance.csv", 
                     header = TRUE, 
                     sep = ";", 
                     skip = 0,
                     row.names = 1)

core <- read.csv("/Documents/Proyectos_R/sentrynels/Data/PPE/Core_ppe_asv_abundance.csv", 
                 header = TRUE, 
                 sep = ";", 
                 skip = 0, 
                 row.names = 1)

ambiente <- read.csv("~/Desktop/sentrynels/Data/Phylo/HCS_Metal_ambienteironment_NMDS.csv", 
                     header = TRUE, 
                     sep = ";", 
                     dec = ".",
                     skip = 0, 
                     row.names = 1)
```

## Check data

```{r}
print(all(colnames(previous)%in%core$Sequences))
```

Como esto dio TRUE, es copiar y pegar el resto de cosas que hice para avanzar... WIIIII
Las fuentes serán Core_Microbiota (y los respectivos markdowns) y los de Metacoder que hice para el ISME

## Data format

```{r}
ambiente$Date <- factor(ambiente$Date, levels = unique(ambiente$Date))
ambiente$Label <- factor(ambiente$Label, levels = unique(ambiente$Label))
ambiente$Site <- factor(ambiente$Site, levels = unique(ambiente$Site))
ambiente$Cu_Level <- factor(ambiente$Cu_Level, levels = unique(ambiente$Cu_Level))

abundance <- core[,-c(1,2,66:81)] %>% t() %>% as.data.frame()
taxa <- core[,-c(3:72)]
write.csv2(taxa, "~/Desktop/sentrynels/Data/Phylo/Taxa_ppe_core.csv")
# I manually checked the file to "fix" the assignation of NA
```

## FASTA file

```{r, echo=FALSE,message=FALSE}
species <- paste(taxa$ASV, taxa$Species, sep = "_")
core_fasta <- data.frame(Names = species, Sequences = taxa$Sequences)
seqRFLP::dataframe2fas(core_fasta, "~/Desktop/sentrynels/Data/Phylo/ppe_core.fasta")
```

Multiple alignment of sequences using the super5 algorithm in MUSCLE v5 (Edgar, year)

## Dendrogram with maximum likelihood for PPE ASV core sequences

```{r}
malign <- "~/Desktop/sentrynels/Data/Phylo/core_multi_aligned.fasta"
dbConn <-dbConnect(SQLite(), ":memory:")
Seqs2DB(malign, type = "FASTA", dbFile = dbConn, "")
x <- dbGetQuery(dbConn, "select description from Seqs")$description
Add2DB(myData = data.frame(identifier = x, stringsAsFactors = FALSE), dbConn)
consensus <- IdConsensus(dbConn, threshold = 0.3, minInformation = 0.1)
distance.matrix <- DistanceMatrix(consensus, correction = "Jukes-Cantor", processors = NULL, verbose = TRUE)
set.seed(420) # Reproducible results
dendro <- TreeLine(myDistMatrix = distance.matrix, 
                   method = "ML", 
                   showPlot = TRUE, 
                   type = "dendrogram", 
                   myXStringSet = consensus, 
                   processors = NULL, 
                   verbose = TRUE)
dbDisconnect(dbConn)
attributes(dendro)
```

## Extras

### Extract colors from ggplot graphs

```{r}
# To simplify my life when exploring data, NOT FINISHED
xtrct <- function(gg.plot) {
  library(ggplot2)
  library(stringr)
  p <- ggplot_build(gg.plot)
  for (i in 1:length(p$data[[1]])) {
    v <- p$data[[1]][[1]]
    c <- str_sub(v, start = 1, end = -3)
    u <- unique(c)
  } 
  print(u)
}
```

